---
title: "Final Project"
author: "Gamze Bilsen, Sierra Cheung, Sehrish Mastoor"
date: "4/30/2022"
output:
  html_document:
    toc: true
    toc_depth: 6
---
```{r,message=FALSE, warning=FALSE,include = FALSE}
library(devtools)
library(dbplyr)
library(dplyr)
library(DT)
library(forcats)
library(geojsonsf)
library(geometries)
library(ggmap)
library(ggplot2)
library(ggrepel)
library(ggthemes)
library(htmlwidgets)
library(leaflet)
library(maps)
library(magrittr)
library(plotly)
library(raster)
library(RColorBrewer)
library(readr)
library(rgdal)
library(rnaturalearth)
library(rnaturalearthdata)
library(rvest)
library(scales)
library(sf)
library(sp)
library(stringr)
library(tidymodels)
library(tidyr)
library(tidytext)
library(tidyverse)
library(tmap)
library(MASS)
library(reshape)
library(reshape2)

#vignette(package = "tmap") # available vignettes in tmap
#vignette("tmap-nutshell")

#devtools::install_github("rstudio/leaflet")
```

```{r,message=FALSE, include=FALSE}
setwd("~/Desktop/Group_J_Climate_Economics/")
finaldata <- read_csv('final_to_use.csv')
df<- read_csv("final_to_use.csv")

#Changing column names ahead of time to match everyone's code
colnames(df)[5] <- 'Total.Deaths'
colnames(df)[6] <- 'Total.Affected'
colnames(df)[68] <- 'perc.Final.Energy.from.Renewables'
colnames(df)[69] <- 'Urban.Population'
colnames(df)[74] <- 'Food.Production'
colnames(df)[82] <- 'Life.Expectancy'

colnames(finaldata)[42] <- 'Arable.Land'
colnames(finaldata)[82] <- 'Life.Expectancy'
```

# Introduction

Climate change is defined as the long term shifts in world temperature and weather patterns. The purpose of this project is to see the different greenhouse gases and climate indicators and the relationship that they have with climate change. We focus on different countries and identify their greenhouse gas emissions to determine which nations are the greatest contributors to the worsening climate crisis.

Climate change is a very real threat to human wellbeing. The changes in our landscapes will accelaratingly affect the wellbeing of many populations through higher instances of natural disasters such as drought or sea level rise. Most of the vulnerable populations are in nations that have historically emitted the least. 

For our project, we're interested in bringing this inequality of current and future well-being to light by looking at changes in climate change indicators such as temperature and freqency of natural disasters and national well-being indicators such as life expectancy and GDP.


# Climate Indicators

## Changes throughout time

We thought the well-documented change that our cognitive dissonance of continuing to go with our day-to-day lives while knowing this accelerating threat to Earth's slow adapting creatures' well-being refuses to fully acknowledge would be a great start.

In the graph above there is a clear increase in temperature from 1980 till 2015, which exhibits a clear correlation with the rise of Co2 Emissions worsening the temperature increases.. When examining average, maximum and minimum temperatures there is a 1 degree celsius  increase in the temperatures. We believe this may surpass the 1.5 degree celsius increase in temperatures before 2025, that were promised by country representatives at the Paris Agreement 
```{r,message=FALSE, warning=FALSE, include=FALSE}

year <- df %>% 
  group_by(Year) %>%
  summarize(sum_death = sum(Total.Deaths,na.rm=TRUE),
            sum_affected = sum(Total.Affected, na.rm=TRUE),
            affected_to_death = sum(Total.Deaths/Total.Affected,na.rm=TRUE),
            `Average Temperature` = mean(tmean,na.rm=TRUE),
            `Maximum Temperature` = mean(tmax,na.rm=TRUE),
            `Minimum Temperature` = mean(tmin,na.rm=TRUE),
            `sum_freq` = sum(Frequency,na.rm=TRUE),
            avg_gdp_cap_growth = mean(gdp_per_cap_growth, na.rm=TRUE),
            sum_no2 = sum(nitrousoxide_emissions,na.rm=TRUE)/1000000,
            CO2 = sum(co2_emission,na.rm=TRUE)/1000000)
year2 <- year[,c('Average Temperature','Maximum Temperature','Minimum Temperature','CO2','Year')]
year3 <- melt(year2, id='Year')
g<-ggplot(year3, aes(x=Year,y=value,color=variable)) + geom_line() +
  xlim(1980,2015) + ggtitle("Overtime Temperature shift and Co2 Emissions") +
  scale_y_continuous(name = "Temperature",
    sec.axis = sec_axis( trans=~.*1000000, name="Co2 Emissions in 100K kT"))  +
    scale_color_manual(values=c('orange','red','light blue','dark blue'),
                       labels=c("Average Temp", "Max Temp", "Min Temp",'Co2 Emissions')) +labs(colour = " ") + 
  theme(legend.position="bottom")

```
```{r, message=FALSE, warning=FALSE}
ggplotly(g)
```

## Emissions and Temperature{.tabset}
Historically the high emitters are plotted in the graphs below. We explore countries that emit both carbon dioxide and nitrous oxide. We focus on the year 2015, as the data for the following years had many missing values. There is a clear trend as China, USA and India are high emitters of the main greenhouse gasses, and due to the size of the countries, economy and population there is a clear contribution made by these economics powers in worsening the climate crisis. There is a clear correlation between high temperatures and high emission of greenhouse gasses for most countries, probably leading to increased frequency of climate related disasters in those countries.
### Temperatures and CO2 emissions in 2015 
```{r,message=FALSE, warning=FALSE, include=FALSE}
## Mean temperature correlation with CO2 emissions for the year 2015 
df1 <- df %>%
    filter(Year == "2015") %>%
    group_by(Country, tmean, co2_emission) %>% 
    summarize(co2_emission) %>%
    arrange(desc(co2_emission)) %>%
    ungroup() %>%
    slice(1:10)
  
## Mean temperature correlation with N.o2 emissions for the year 2015 
df2 <- df %>%
    filter(Year == "2015") %>%
    group_by(Country, tmean, nitrousoxide_emissions) %>% 
    summarize(nitrousoxide_emissions) %>%
    arrange(desc(nitrousoxide_emissions)) %>%
    ungroup() %>%
    slice(1:10)


t1 <- plot_ly(
  data = df1,
  x = ~tmean,
  y = ~co2_emission,
  legendtitle=TRUE,
  type = "scatter",
  mode = "markers",
  color = ~Country,
  colors = 'RdBu',
  margin = 20,
  marker = list(
                size=20,
                opacity=.9)
  ) %>% 
  layout( title = 'Highest Co2 Emissions with Mean Temperature',xaxis = list(title = 'Mean Temperature'), yaxis = list(title = 'Co2 Emissions'))

t1 <- t1 %>% layout(legend=list(title=list(text='<b> Countries </b>')))
```
```{r, message=FALSE, warning=FALSE}
t1
```

### Mean Temperatures and Nitrous Oxide
```{r,message=FALSE, warning=FALSE, include = FALSE}
x <- plot_ly(
  data = df2,
  x = ~tmean,
  y = ~nitrousoxide_emissions,
  type = "scatter",
  mode = "markers",
  color = ~Country,
  colors = 'RdYlBu',
  marker = list(
                size=20,
                opacity=.9)
  ) %>% 
  layout(title = 'Highest Nitrous Oxide Emissions with Mean Temperature',xaxis = list(title = 'Mean Temperature'),
         yaxis = list(title = 'Nitrous Oxide Emissions'))
   
x <- x %>% layout(legend=list(title=list(text='<b> Countries </b>')))
```
```{r, message=FALSE, warning=FALSE}
x
```



# Climate Disasters

## Countries affected most by climate related disasters based on total number of deaths from 2010 to 2015

Countries affected most by climate related disasters based on total number of deaths from 2010 to 2015

```{r, message=FALSE, warning=FALSE, include=FALSE}
labels <- c("# Disasters" , "# Deaths" , "# Affected")
colors <- c('orange','red','purple')
temp_to_dis <- df %>% 
  group_by(tmean) %>%
  summarize(num_disaster = sum(Frequency,na.rm=TRUE),
            total_deaths = sum(Total.Deaths,na.rm=TRUE)/1000,
            total_affected = sum(Total.Affected,na.rm=TRUE)/1000000)
molten.temp <- melt(data.frame(temp_to_dis), id = c("tmean"))
mtemp <- ggplot(molten.temp,aes(x=tmean,y=value,color=variable)) + 
  geom_smooth(method = lm, se = FALSE) +
  scale_color_manual(name = "Y series", labels = labels, 
                     values = colors) + 
  ggtitle("Temperature vs Disaster Effects")
```
```{r,message=FALSE, warning=FALSE}
ggplotly(mtemp)
```



### Climate Disasters and Death Counts
We have graphed total climate disasters according to years. We focus on 2010 till 2015, and the graphed results show that developing nations are most impacted due to the climate change. This aligns with the fact that developing nations usually are the worst impacted by actions of the developed nations. China and India are present in all the graphs, whereas the United States is present in 4 of them. These results show that lower emitters are clearly impacted by the actions of higher emitters, although higher emitters are also seeing the negative results of their economic growth.

```{r, message=FALSE, warning=FALSE, include=FALSE}
require(gridExtra)
death <- df %>%
  filter(Year>=2010&Year<=2015) %>%
    group_by(Country,Year) %>%
    summarize(total_deaths =sum(Total.Deaths,na.rm=TRUE)) %>%
    arrange(desc(total_deaths))

ten<-death %>% filter(Year==2010) %>% arrange(desc(total_deaths)) %>%ungroup()%>%slice(1:10)
eleven<-death %>% filter(Year==2011) %>% arrange(desc(total_deaths)) %>%ungroup()%>%slice(1:10)
twelve<-death %>% filter(Year==2012) %>% arrange(desc(total_deaths)) %>%ungroup()%>%slice(1:10)
thirt<-death %>% filter(Year==2013) %>% arrange(desc(total_deaths)) %>%ungroup()%>%slice(1:10)
fourt<-death %>% filter(Year==2014) %>% arrange(desc(total_deaths)) %>%ungroup()%>%slice(1:10)
five<-death %>% filter(Year==2015) %>% arrange(desc(total_deaths)) %>%ungroup()%>%slice(1:10)

ten$total_deaths<-ten$total_deaths/1000
eleven$total_deaths<-eleven$total_deaths/1000
twelve$total_deaths<-twelve$total_deaths/1000
thirt$total_deaths<-thirt$total_deaths/1000
fourt$total_deaths<-fourt$total_deaths/1000
five$total_deaths<-five$total_deaths/1000

plot1 <- ggplot(ten, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab('Countries') +ylab(' ') +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size=10),
        title = element_text(size=10))+ylim(0,21) + ggtitle('2010') 
plot2 <- ggplot(eleven, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab(' ') +ylab(' ')  +
  theme(title = element_text(size=10)) + ggtitle('2011')
plot3 <- ggplot(twelve, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab(' ') +ylab(' ')+ylim(0,21)  +
  theme(title = element_text(size=10)) + ggtitle('2012')
plot4 <- ggplot(thirt, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab('Countries') + ylab('Total Deaths in 1K') +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size=10),
        title = element_text(size=10))+ylim(0,21) + ggtitle('2013')
plot5 <- ggplot(fourt, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab(' ') +ylab('Total Deaths in 1K') +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size=10),
        title = element_text(size=10))+ylim(0,21)+ ggtitle('2014')
plot6 <- ggplot(five, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab(' ') +ylab('Total Deaths in 1K') +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size=10),
        title = element_text(size=10)) +ylim(0,21)+ ggtitle('2015')

```
```{r, messge=TRUE, warning=FALSE}
grid.arrange(plot1, plot2, plot3,plot4,plot5,plot6, ncol=3,nrow=2, 
             top='Top 10 Climate Disasters by Total Number of Deaths in 1K per Year')
```



### Worldwide Temperatures and Natural Disaster Deaths

The Map below shows the temperature averages alongside the max and minimum to show temperature ranges between the year.  The pop up also displays total deaths from disasters. The graph is coloured according to the average temperature to identify whether there is really a relationship between temperatures and total deaths. 
```{r,message=FALSE, warning=FALSE, include=FALSE}
countryfinal <- df %>% 
  filter(Year>=2010&Year<=2015)%>%
  group_by(Country) %>%
  summarize(sum_death = sum(Total.Deaths,na.rm=TRUE),
            sum_affected = sum(Total.Affected, na.rm=TRUE),
            affected_to_death = sum(Total.Deaths/Total.Affected,na.rm=TRUE),
            average_temp = mean(tmean,na.rm=TRUE),
            sum_freq = sum(Frequency,na.rm=TRUE),
            avg_gdp_cap_growth = mean(gdp_per_cap_growth, na.rm=TRUE),
            average_perc_urban = mean(Urban.Population, na.rm=TRUE),
            avg_gdp = mean(GDP),
            avg_gdp_cap = mean(gdp_per_cap,na.rm=TRUE),
            avg_temp = mean(tmean,na.rm=TRUE),
            average_perc_arable = mean(perc_arable_land_area, na.rm=TRUE),
            average_food_prod = mean(Food.Production, na.rm=TRUE),
            average_pop = mean(Population,na.rm=TRUE),
            avg_life_exp = mean(Life.Expectancy,na.rm=TRUE),
            avg_perc_forest = mean(perc_forest_land_area,na.rm=TRUE),
            avg_perc_energy_ren = mean(perc.Final.Energy.from.Renewables,na.rm=TRUE),
            avg_co2 = mean(co2_emission,na.rm=TRUE),
            avg_remittances = mean(remittances_received,na.rm=TRUE),
            avg_elect_days = mean(days_for_electricity,na.rm=TRUE),
            average_max_temp = mean(tmax,na.rm=TRUE),
            average_min_temp = mean(tmin,na.rm=TRUE),
            average_rural_pop = mean(`Percent Rural Pop`,na.rm=TRUE),
            average_urban_pop = mean(`% Urban Population`,na.rm=TRUE))

world <- ne_countries(scale = "medium", returnclass = "sf")
world_map <- merge(world, countryfinal, by.x="name", by.y="Country")

population_data<-na.omit(df[,c('Country','Year','Population')])%>% 
  group_by(Country) %>%
  slice(which.max(Year))

world_map <- merge(world_map,population_data,by.x='name',by.y='Country')

labels <- paste(
  "<strong>",world_map$name,"</strong><br/>Average Temperature:",round(world_map$avg_temp,2),"C",
  "<br/>Maximum Temperature:",round(world_map$average_max_temp,2), "C",
  "<br/>Minimum Temperature:",round(world_map$average_min_temp,2), "C",
  "<br/>Total Deaths from Disasters:",round(world_map$sum_freq,2)
) %>% lapply(htmltools::HTML)

```
```{r, message=FALSE, warning=FALSE}
leaflet(world_map)%>%
 addPolygons(stroke = TRUE, smoothFactor = 0.5,
  weight=1, color='#333333', opacity=1, 
  fillColor = ~colorQuantile("Blues", avg_temp)(avg_temp), 
  fillOpacity = 1,
  highlightOptions = highlightOptions(
    weight = 3,
    fillOpacity = 0.7,
    bringToFront = TRUE),
  popup = labels)
```



```{r,message=FALSE, warning=FALSE, include=FALSE}
world_map1 <- merge(world, countryfinal, by.x="name", by.y="Country")

population_data1<-na.omit(df[,c('Country','Year','Population')])%>% 
  group_by(Country) %>%
  slice(which.max(Year))

world_map1 <- merge(world_map1,population_data1,by.x='name',by.y='Country')

labels <- paste(
  "<strong>",world_map1$name,"</strong><br/>Average GDP per capita:","$",round(world_map1$avg_gdp_cap,2),
  "<br/>Urban Population:",round(world_map1$average_urban_pop,2), "%",
  "<br/>Rural Population:",round(world_map1$average_rural_pop,2), "%"
) %>% lapply(htmltools::HTML)
```


## CO2 Emitter Rankings Worldwide

The map below shows 10 of the top and bottom CO2 emitters for they year the 2015. We can see that most of the highest emitters are economic powerhouses such as China and United States, and the lower emitters tend to mostly be island nations. We can also find the life expectancy and GDP for each country.
```{r,message=FALSE, warning=FALSE, include=FALSE}
top10co2 <- filter(finaldata, Year ==2015)%>%
  group_by(Year, Country, co2_emission, Life.Expectancy, GDP)%>%
  summarize(count_co2=co2_emission, count_life=Life.Expectancy, count_gdp =GDP) %>%
  arrange(desc(co2_emission))%>%
  ungroup()%>%
  slice(1:10)

bottom10co2 <- filter(finaldata, Year ==2015)%>%
  group_by(Year, Country, co2_emission, Life.Expectancy, GDP)%>%
  summarize(count_co2=co2_emission, count_life=Life.Expectancy, count_gdp =GDP) %>%
  arrange(desc(-co2_emission))%>%
  ungroup()%>%
  slice(1:10)

top10co2['rank'] <- 'Top 10 Emitter'
bottom10co2['rank'] <- 'Bottom 10 Emitter'


#binding top and bottom 10 co2 level countries
top10_bottom10 <- rbind(top10co2, bottom10co2)

#Longitude and latitudes for top and bottom 10 countries
longitude <- c(116.20, #China
               -77.02, #US
               77.13, #India
               138.25, #Japan
               13.25, #Germany
               45.00, #Saudi Arabia
               -106.42, #Canada
               113.10, #Indonesia
               -51.92, #Brazil
               -102.55, #Mexico
               177.65, #tuvalu
               166.3, #Nauru
               -168.73, #Kiribati
               -175.20, #Tonga
               6.61, #Sao Tome and Principe
               173, #Vanuatu
               -171.00, #Marshall Islands
               43.33, #Comoros
               9.55, #Liechtenstein
               -61.37) #Dominica

latitude <- c(39.55, #China
              39.91, #US
              28.37, #India
              36.20, #Japan
              52.30, #Germany
              23.88, #saudi Arabia
              56.27, #Canada 
              -6.09, #Indonesia
              -14.47, #Brazil
              23.63, #Mexico
              -7.10, #tuvalu
              -0.523, #Nauru
              -3.37, #Kiribati
              -21.10, #Tonga
              0.16, #Sao Tome and Principe
              -17.45, #Vanuatu
              7.13, #Marshall Islands
              -11.64, #Comoros
              47.08, #Liechtenstein
              15.41) #Dominica

top10_bottom10['longitude'] <- longitude
top10_bottom10['latitude'] <- latitude

#adding pop-up contents
popup_content <- paste("Country:",top10_bottom10$Country,"<br/>",
                       "CO2 Emissions (Hundred Thou. KT):",(top10_bottom10$co2_emission)*1000,"<br/>",
                       "2015 Life Expetancy (Years):",round(top10_bottom10$Life.Expectancy, digits = 3),"<br/>",
                       "2015 GDP (in Billions):",top10_bottom10$GDP,"<br/>")
#assigning colors to countries
pal = colorFactor("RdBu", domain = top10_bottom10$rank)
color_avail = pal(top10_bottom10$rank)

```
```{r, message=FALSE, warning=FALSE}
#plotting on world map
leaflet(top10_bottom10) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~longitude, lat = ~latitude, group = "Country",
                   fillColor=color_avail, stroke = TRUE, fillOpacity = 1, popup = popup_content)%>%
  addLegend(pal = pal, values = ~top10_bottom10$rank, title = "CO2 Emitters")
```

# GHG Emissions
## Carbon dioxide Emissions from 2010 till 2015

The plot below shows the exact levels of CO2 emissions of the top 10 emitters summed up from years 2010 to 2015. We can see a significant difference between the top three emitters, with China being significantly above US, and the US being significantly above India, then the emissions for each country slowly stalling off. Thus, there seems to be an acceleration to emission per each country given driving factors that aren’t included in this map, however could be attributed to current and historic economic growth levels per capita.
```{r,message=FALSE, warning=FALSE, include=FALSE}
# Countries with the total highest Co2 Emissions from 2010 to 2015 
co2 <- df %>%
    filter(Year>="2010" , Year <="2015") %>%
    group_by(Year, Country, co2_emission)

totalco2 <- co2 %>% 
    group_by(Year,Country, co2_emission) %>%
    group_by(Country) %>%
    summarize(Count = sum(co2_emission)) %>% 
    arrange(desc(Count)) %>%
    ungroup() %>%
    slice(1:10)

p1 <- plot_ly(
  data = totalco2,
  x = ~Count,
  y = ~Country,
  legendtitle=TRUE,
  type = "scatter",
  mode = "markers",
  color = ~Country,
  colors = 'RdBu',
  margin = 20,
  marker = list(
                size=20,
                opacity=.9)
  ) %>% 
layout(yaxis = list(categoryorder = "total ascending")) %>% 
  layout( title = 'Countries Carbon dioxide Emissions from 2010 till 2015',xaxis = list(title = 'Carbon dioxide levels'), yaxis = list(title = 'Countries '))

p1 <- p1 %>% layout(legend=list(title=list(text='<b> Countries </b>')))

```
```{r, message=FALSE, warning=FALSE}
p1
```
## Country Wellbeing Plots

### Arable Land for Top and Bottom CO2 Emitters

Arable Land for Top and Bottom 10 CO2 Emitters
The plot below shows countries that have less arable land tend to have emitted lower emissions except for China which despite its massive land area, they likely have forest and mountainous regions that cover most of the landmass. United States is the only country that seems to be balanced with being in the middle of the extreme values for both variables.
```{r,message=FALSE, warning=FALSE, include=FALSE}
dataco2 <- filter(finaldata, is.na('co2_emission') == FALSE)%>%
  filter(Year > 2009, Year < 2016)%>%
  group_by(Year, Country, co2_emission)

dataco2 <- filter(finaldata, Year == 2015)%>%
  group_by(Country, co2_emission, Arable.Land) %>%
  summarize(count_co2=co2_emission, count_ar=Arable.Land) %>%
  arrange(desc(co2_emission))%>%
  ungroup()%>%
  slice(1:10)

#plotting co2 and arable land

plot1  <- plot_ly(dataco2, x = ~count_co2, y = ~count_ar, 
                 type = 'scatter',   color = ~Country,
  colors = 'RdYlBu', marker = list( size=15,opacity=.9)) %>% 
  layout( title = 'Countries Carbon dioxide Emissions from 2010 till 2015',xaxis = list(title = 'Average CO2 Emissions (Millions of Kilotons)'), yaxis = list(title = "Average Arable Land and CO2 Emissions from 2010-2015"))

plot1  <- plot1   %>% layout(legend=list(title=list(text='<b> Countries </b>')))

```
```{r, message=FALSE, warning=FALSE}
plot1
```



### Arable Land and Percentage of Forested Land
In the plot below we can explore the realtionship between arable and forested lands and we can see that countries with higher forested land have lower arable lands.
```{r,message=FALSE, warning=FALSE, include=FALSE}
arableland_gdp<- ggplot(countryfinal,aes(y=avg_perc_forest,x=average_perc_arable, color=avg_perc_forest)) + 
  geom_smooth(method = lm, se = FALSE) + 
  geom_point(aes(size=avg_gdp)) + 
  ggtitle("Average Percent Arable Land to Average Percent Forest Area per Country of Years 2010 to 2015 and GDP") +
  xlab("Average Arable Land Percent per Country") + 
  ylab('Average Forest Land Percent per Country') +
  theme_minimal() +
  labs(size='Average GDP')
```
```{r,message=FALSE, warning=FALSE}
ggplotly(arableland_gdp)
```

### Economic and Population Wellbeing and CO2 Emissions
Co2 emissions + GDP (how top 10 and bottom 10 GDP countries compare in CO2 emissions from 2010-2015)

Below is a plot of the top and bottom CO2 Emitters worldwide. 
```{r,message=FALSE, warning=FALSE, include=FALSE}
library(plotly)
plot3 <- plot_ly(top10_bottom10, x = ~Country, y = ~Life.Expectancy, 
                 type = 'bar', color = ~rank, colors = "Reds") %>%
  layout(xaxis = list(categoryorder = "total ascending")) %>% 
  layout( title = 'Top and Bottom Co2 Emitters Worldwide according to Life Expectancy',xaxis = list(title = 'Country'), yaxis = list(title = "Life Expectancy"))

plot3 <- plot3  %>% layout(legend=list(title=list(text='<b> Category </b>')))
```
```{r, message=FALSE, warning=FALSE}
plot3
```


### GDP of countries Worldwide in 2015

The map below shows the top 10 countries based on GDP. This is another form of understanding of the wellbeing of countries, assuming higher GDP countries’ citizens have improved wellbeing. We can recognise most of the names from higher CO2 emitters as well such as countries like US and China.
```{r,message=FALSE, warning=FALSE, include=FALSE}
## GDP for the year 2020 

gdp <- df %>%
  filter(Year == "2015") %>%
    group_by(Country, GDP) %>%
    #summarize(count=GDP) %>%
    arrange(desc(GDP))

# Top 10 Countries according to GDP
top10gdp <-gdp %>% 
ungroup() %>%
slice(1:10)

plot5  <- plot_ly(top10gdp, y = ~Country, x = ~GDP, orientation = 'h',
                  type = 'bar', color = ~Country, colors = 'RdBu') %>% 
layout(yaxis = list(categoryorder = "total descending"),showlegend = FALSE) %>% 
  layout( title = 'Top 10 Countries according to GDP',xaxis = list(title = 'GDP'), yaxis = list(title = "Countries"))

#plot5 <- plot5  %>% layout((showlegend = FALSE))
```
```{r, message=FALSE, warning=FALSE}
plot5  
```


The Map below shows the rural and urban population percentages alongside the GDP per capita. The graph is coloured according to GDP per capita which is a direct indicator of a countries economic well being. We link this to rural and urban populations to see whether there is a connection. 
```{r, message=FALSE, warning=FALSE}
leaflet(world_map1)%>%
 addPolygons(stroke = TRUE, smoothFactor = 0.5,
  weight=1, color='#333333', opacity=1, 
  fillColor = ~colorQuantile("Reds", avg_gdp_cap)(avg_gdp_cap), 
  fillOpacity = 1,
  highlightOptions = highlightOptions(
    weight = 3,
    fillOpacity = 0.7,
    bringToFront = TRUE),
  popup = labels)
```



### Deaths in Countries Worldwide in 2015
```{r,message=FALSE, warning=FALSE, include=FALSE}
## Total deaths for the year 2020 
death <- df %>%
  filter(Year == "2015") %>%
    group_by(Country, Total.Deaths) %>%
    #summarize(count = Total.Deaths) %>%
    arrange(desc(Total.Deaths))

# Top 10 Countries according to death 
topdeath <-death %>% 
ungroup() %>%
slice(1:10)


plot6  <- plot_ly(topdeath, y = ~Country, x = ~Total.Deaths, orientation = 'h',
                  type = 'bar', color = ~Country, colors = 'RdBu') %>% 
layout(yaxis = list(categoryorder = "total descending"),showlegend = FALSE) %>% 
  layout( title = 'Top 10 Countries according to Total Deaths: 2015',xaxis = list(title = 'Total Deaths'), yaxis = list(title = "Countries"))

plot6 <- plot6  %>% layout((showlegend = FALSE))
```
```{r, message=FALSE, warning=FALSE}
plot6  
```


### Focus on Vital Indicators

The map below shows the relationship between average temperature, average GDP, total disasters and total death count. There is a clear correlation between lower average GDP and higher temperatures and higher death count. The line on the plot shows to us that higher GDP nations have lower temperature, reinforcing the fact that the negative impact of climate is distributed amongst the world, whereas the benefits may not be.
```{r,message=FALSE, warning=FALSE, include=FALSE}
# GDP per capita with total deaths Frequency 

df1 <- df[, c('Country', 'Year', 'gdp_per_cap', 'Total.Deaths','Frequency','tmean')]

x <- df1 %>%
  filter(Year>=2010 , Year <=2015) %>%
  #group_by(gdp_per_cap, Total Deaths,Frequency) %>% 
  group_by(Country) %>%
  summarize('Total.Death.Count' = sum(`Total.Deaths`, na.rm=TRUE),
            'Total Frequency' = sum(Frequency, na.rm=TRUE),
            'Average GDP' = mean(gdp_per_cap, na.rm=TRUE),
            'Average Temperature' = mean(tmean,na.rm=TRUE)
            )

s<- ggplot(x,aes(y=`Average GDP`,x= `Average Temperature`)) + 
  geom_smooth(method = lm, se = FALSE) + 
  geom_point(aes(size=`Total Frequency`, colour= `Total.Death.Count`, alpha=14)) + 
  ggtitle("Relationship between GDP and Temperature") + xlab("Mean Temperature") +  ylab('Average GDP') +
  theme_minimal() +
  labs(size='Total.Death.Count')
```
```{r, message=FALSE, warning=FALSE}
ggplotly(s)
```

# Thoughts and Insights
Data Collection in different countries differ so how valuable is all the data when comparing to developing nations and developed.
There is a clear inequity between emitters and the impacted nations, so having more valuable data would help us delve into this deeper.









