---
title: "Final Project"
author: "Gamze Bilsen"
date: "4/30/2022"
output: html_document
---

```{r}
library(dplyr)
library(tidyr)
library(tidytext)
library(ggplot2)
library(DT)
library(readr)
library(raster)
library(htmlwidgets)
library(rgdal)
library(sf)
library(ggmap)
library(tmap)
library(maps)
library(DT)
library(magrittr)
library(dplyr)
library(ggthemes)
library(plotly)
library(DT)
library(stringr)
library(plotly)
library(RColorBrewer)
library(scales)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(leaflet)
```

# Climate Indicators

Decrease in weather related disasters: https://www.nationalgeographic.com/environment/article/weather-disaster-related-deaths-are-downwarming-could-undo-that-trend

## Change throughout time
```{r}
year <- df %>% 
  group_by(Year) %>%
  summarize(sum_death = sum(Total.Deaths,na.rm=TRUE),
            sum_affected = sum(Total.Affected, na.rm=TRUE),
            affected_to_death = sum(Total.Deaths/Total.Affected,na.rm=TRUE),
            average_temp = mean(tmean,na.rm=TRUE),
            average_max_temp = mean(tmax,na.rm=TRUE),
            average_min_temp = mean(tmin,na.rm=TRUE),
            sum_freq = sum(Frequency,na.rm=TRUE),
            avg_gdp_cap_growth = mean(gdp_per_cap_growth, na.rm=TRUE),
            sum_no2 = sum(nitrousoxide_emissions,na.rm=TRUE)/1000000,
            sum_co2 = sum(co2_emission,na.rm=TRUE)/1000000)
ggplot(year, aes(x=Year)) +
  geom_line(aes(y=average_temp)) +
  geom_line(aes(y=sum_co2)) +
  geom_line(aes(y=average_max_temp),color='red') +
  geom_line(aes(y=average_min_temp),color='blue') + 
  xlim(1980,2015) +
  scale_y_continuous(name = "Temperature",
    sec.axis = sec_axis( trans=~.*1000000, name="CO2 Emissions"))
  
```

## Country Level Information
```{r}
countryfinal <- df %>% 
  filter(Year>=2010&Year<=2015)%>%
  group_by(Country) %>%
  summarize(sum_death = sum(Total.Deaths,na.rm=TRUE),
            sum_affected = sum(Total.Affected, na.rm=TRUE),
            affected_to_death = sum(Total.Deaths/Total.Affected,na.rm=TRUE),
            average_temp = mean(tmean,na.rm=TRUE),
            sum_freq = sum(Frequency,na.rm=TRUE),
            avg_gdp_cap_growth = mean(gdp_per_cap_growth, na.rm=TRUE),
            average_perc_urban = mean(X..Urban.Population,na.rm=TRUE),
            avg_gdp = mean(GDP),
            avg_gdp_cap = mean(gdp_per_cap,na.rm=TRUE),
            avg_temp = mean(tmean,na.rm=TRUE),
            average_perc_arable = mean(perc_arable_land_area, na.rm=TRUE),
            average_food_prod = mean(Food.Production, na.rm=TRUE),
            average_pop = mean(Population,na.rm=TRUE),
            avg_life_exp = mean(Life.Expectancy,na.rm=TRUE),
            avg_perc_forest = mean(perc_forest_land_area,na.rm=TRUE),
            avg_perc_energy_ren = mean(X..Final.Energy.from.Renewables,na.rm=TRUE),
            avg_co2 = mean(co2_emission,na.rm=TRUE),
            avg_remittances = mean(remittances_received,na.rm=TRUE),
            avg_elect_days = mean(days_for_electricity,na.rm=TRUE),
            average_max_temp = mean(tmax,na.rm=TRUE),
            average_min_temp = mean(tmin,na.rm=TRUE))

world <- ne_countries(scale = "medium", returnclass = "sf")
world_map <- merge(world, countryfinal, by.x="name", by.y="Country")

population_data<-na.omit(df[,c('Country','Year','Population')])%>% 
  group_by(Country) %>%
  slice(which.max(Year))

world_map <- merge(world_map,population_data,by.x='name',by.y='Country')

labels <- paste(
  "<strong>",world_map$name,"</strong><br/>Average Temperature",round(world_map$avg_temp,2),"C",
  "<br/>Maximum Temperature",round(world_map$average_max_temp,2), "C",
  "<br/>Minimum Temperature",round(world_map$average_min_temp,2), "C",
  "<br/>Total Deaths from Disasters",round(world_map$sum_freq,2)
) %>% lapply(htmltools::HTML)

leaflet(world_map)%>%
 addPolygons(stroke = TRUE, smoothFactor = 0.5,
  weight=1, color='#333333', opacity=1, 
  fillColor = ~colorQuantile("Greens", avg_temp)(avg_temp), 
  fillOpacity = 1,
  highlightOptions = highlightOptions(
    weight = 3,
    fillOpacity = 0.7,
    bringToFront = TRUE),
  popup = labels)
```

# Climate Disasters

```{r}
df1 <- df[, c('Country', 'Year', 'gdp_per_cap', 'Total.Deaths','Frequency')]

x <- df1 %>%
  filter(Year>="2010" , Year <="2015") %>%
 group_by(Year) %>%
  summarize(totaldeathcount = sum(Total.Deaths, na.rm=TRUE),
            totalfrequency = sum(Frequency, na.rm=TRUE),
            avggdp = mean(gdp_per_cap, na.rm=TRUE)
            )

```

# Climate Disasters
## Total amount of disasters and total deaths throughout the last 20 years

```{r}
ggplot(year, aes(x=Year)) +
  geom_line(aes(y=average_temp)) +
  geom_line(aes(y=sum_co2)) +
  geom_line(aes(y=average_max_temp),color='red') +
  geom_line(aes(y=average_min_temp),color='blue') + 
  xlim(1980,2015) +
  scale_y_continuous(name = "Temperature",
    sec.axis = sec_axis( trans=~.*1000000, name="CO2 Emissions"))
```


## Countries affected most by climate related disasters based on total number of deaths from 2010 to 2015
```{r}
require(gridExtra)
death <- df %>%
  filter(Year>=2010&Year<=2015) %>%
    group_by(Country,Year) %>%
    summarize(total_deaths =sum(Total.Deaths,na.rm=TRUE)) %>%
    arrange(desc(total_deaths))


plot1 <- ggplot(ten, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab('Countries') +ylab(' ') +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size=10))
plot2 <- ggplot(eleven, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab(' ') +ylab(' ')
plot3 <- ggplot(twelve, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab(' ') +ylab(' ')
plot4 <- ggplot(thirt, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab('Countries') +ylab('Total Deaths') +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size=10))
plot5 <- ggplot(fourt, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab(' ') +ylab('Total Deaths') +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size=10))
plot6 <- ggplot(five, aes(y=total_deaths, x=reorder(Country, desc(total_deaths), fill = Country, color_continuous_scale=scale))) + 
  geom_col() + coord_flip() +xlab(' ') +ylab('Total Deaths') +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size=10))

grid.arrange(plot1, plot2, plot3,plot4,plot5,plot6, ncol=3,nrow=2, 
             top='Top 10 Climate Disasters by Total Number of Deaths per Year') 

```

# GHG Emissions



# Wellbeing

## Forest Land Area to Arable Land used for Agriculture Area per Country Based on GDP
```{r}
ggplot(countryfinal,aes(y=avg_perc_forest,x=average_perc_arable)) + 
  geom_smooth(method = lm, se = FALSE) + 
  geom_point(aes(size=avg_gdp)) + 
  ggtitle("Average Percent Arable Land to Average Percent Forest Area per Country of Years 2010 to 2015 and GDP") +
  xlab("Average Arable Land Percent per Country") + 
  ylab('Average Forest Land Percent per Country') +
  theme_minimal() +
  labs(size='Average GDP')
```

# Climate Indicators and Disasters
```{r}
temp_to_dis <- df %>% 
  group_by(tmean) %>%
  summarize(num_disaster = sum(Frequency,na.rm=TRUE),
            total_deaths = sum(Total.Deaths,na.rm=TRUE),
            total_affected = sum(Total.Affected,na.rm=TRUE))
ggplot(temp_to_dis,aes(x=tmean)) + 
  geom_smooth(aes(y=num_disaster),method = lm, se = FALSE) #+
  #geom_smooth(aes(y=total_deaths),method = lm, se = FALSE)
```

# Climate Disasters and Wellbeing

# Climate Indicators and Wellbeing












